
\documentclass[subfig,blackref,approvalform]{drexel-thesis}

%% Enter the appropriate information here
\author{Vincent J. Baker}
\title{Traveling waves in low-dimensional neuronal systems}
\DUTmonth{September}
\DUTyear{2021}
\degree{Doctor of Philosophy}
\advisor{Luis Cruz, Ph. D.}
\copyrighttext{\copyrighttextCCBYSA}

\usepackage[super,sort&compress,comma]{natbib} % fancy citation extensions
\bibliographystyle{unsrtnat}
%\bibliographystyle{ieeetr}
\usepackage{notoccite}

\usepackage{fancyvrb} % nicer verbatim handling
\DefineShortVerb{\|}  % \verb+TEXT+  ->  |TEXT|

\renewenvironment{preamble}{% look, we're clobbering an environment!
  What preamble?\begingroup\color{white}}{\endgroup}

\usepackage[english]{babel} % English blindtext
\usepackage{blindtext} % macros for producing dummy output
\addto\captionsenglish{\DUTnames}

%%Begin packages used in previous papers
\usepackage{amsmath}
\usepackage{times}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{wrapfig}
\usepackage{csquotes}
\usepackage{xcolor}
\usepackage{rotating}
\usepackage{bbm}
\usepackage{latexsym}
\usepackage{placeins}


\begin{document}
\begin{DUTpreamble} % "DUT" because original preamble was clobbered.

\begin{dedications}
\rule{\textwidth}{1pt}
\\
\centering
\textsc{This thesis is dedicated to my wife Amanda and my mother Michelle.
Their unwavering support and encouragement kept me going when it would have been all too easy to abandon this work.}
\\
\rule{\textwidth}{1pt}
\end{dedications}


\begin{acknowledgments}
I would first like to thank my advisor Dr. Luis Cruz. 
His patient guidance kept me coming back to the path, despite my tendency to wander.
I'd like to thank mt commitee,  Drs. Urbanc, Richards, Liang and Rosenwinkel, for their patience with my early efforts in finding the thread of this research.
Thanks to my colleagues at Lockheed Martin's Advanced Technology Laboratories, especially the management team who allowed me to pursue my education in parallel with my professional requirements.

Finaly I would like to thank my son Alex. 
He was 9 months old when I started this journey, and I'm thrilled he got to see me finish it.
\end{acknowledgments}

\tableofcontents
\listoftables
\listoffigures

\begin{abstract}
Traveling waves of neuronal activity have been observed in many locations within the mammalian cortex.
Several functional roles have been proposed for these traveling waves including visual and audio processing, information transfer between cortical areas and memory consolidation.
Computer simulations of such traveling waves are typically performed in strictly one-dimensional or two-dimensional systems using homogeneous neurons with isotropic connectivity. 
Here we perform a computational investigation of traveling waves in more diverse systems with variations in neuron parameters and connectivity.
We explore quasi one-dimensional and quasi two-dimensional networks of heterogeneous neurons with a biologically influenced computational model of neuron dynamics and connectivity.
Our simulations reveal the model parameters for which traveling waves are supported, as well as how these parameters affect wave characteristics such as speed.
We find that a uniform random stimulus reliably evokes traveling waves in networks with local connectivity between neurons, with low-threshold-spiking inhibitory neurons influencing the wave initiation sites.
We reproduce the circular, plane and spiral wave patterns that have been observed in vivo and demonstrate the spiral waves are mroe likely to form when the inter-neuron connections are weaker.
We extend previous work suggesting that traveling waves mediate information transfer by demonstrating that traveling waves can act as an encoding and transport mechanism for neural population activity.

\end{abstract}
\end{DUTpreamble}

\begin{thesis}
\pdfbookmark[-1]{Main Matter}{Main Matter}

\include{c1/Introduction}
\include{c2/Methods}
\include{c3/quasi1D}
\include{c4/lowdim_systems}
\include{c5/EncodingTransport}
\include{c6/Conclusion}
\end{thesis}

\bibliography{thesis}

%\appendix
%\include{}

%\begin{vita}

%\end{vita}

\end{document}
\endinput
%%
%% End of file `example.tex'.
